---
title: 'EDA: Housing Dominican Republic'
output:
  html_document:
    df_print: paged
---

# Introduction

For this analysis, we will examine the effect that area (measured in squared meters) has on the value of apartment, for residential use, in Santo Domingo, Domincan Republic. Prices and apartment's characteristics are collect via web scraping. Specifically, data were retrieved on 22nd of march, 2022, from supercasas.com, a beacon on the online dominican real estate market.

# Loading libraries and data

```{r loading-libraries, message = FALSE}
rm(list = ls())

options(scipen = 999)

library(robustbase)
library(tidyverse)
library(caret)

set.seed(1234)
```

Data on every listing available on several dates at supercasas.com were retrieved. Here, we load and filter said data so we only have information on apartments for residential use on the date of interest. Then, we split the dataset into training (70%) and testing (30%).

```{r importing-data}
path <- "./1_data/0_raw/housing price/"
housing_files <- list.files(path)
housing <- read_csv(paste0(path, housing_files))

housing <- housing %>%
  filter(date ==  "2022-03-22",
         usage == "Residencial",
         province %in% c("Santo Domingo", "Santo Domingo Centro (D.N.)")) %>%
  rename(location = neighborhood) %>%
  select(-c(date, usage, city, province))

inTrain <- createDataPartition(housing$price.usd, p = 0.7, list = FALSE)

training <- housing[inTrain, ]
testing <- housing[-inTrain, ]
```

# Data cleaning

First, let's see what's on the dataset:

```{r}
glimpse(training)
```
Our training dataset contains `r nrow(training) %>% format(big.mark = ",")` observations and `r ncol(training) %>% format(big.mark = ",")` variables. Of them, we can highlight `price`, `currency` and `price.usd`. `price` and `currency` is the actual price shown in the listing's site. They can be in local currency or US dollars, depending on the seller's preference. `price.usd` is a user-made feature of prices in US dollars. Hence, if prices were stated in local currency, they were converted into US dollars. Otherwise, they stay the same. We'll drop `currency` and `price` and keep `price.usd`. Then, and for simplicity's sake, we'll rename `price.usd` as `price`.

```{r}
training <- training %>%
  select(-c(currency, price)) %>%
  rename(price = price.usd)
```

When looking at the proportion of `NA`s are present per variable, over 50% of listings did not provide information regarding the floor the apartment is located at. The proportion of missing values for all other variables is acceptable. And so, we removed that variable.

```{r}
apply(training, 2, 
      \(x) {
        n <- length(x)
        na <- x %>%
          is.na() %>%
          sum()
        prop.na <- na / n * 100
        })

training <- training %>%
  select(-c(story))
```

Before implementing some formal procedure for outlier removal, let's analyse the data at hand. First, we can see that `area` has some outstanding observations: a minimum value of `r min(training$area, na.rm = TRUE) %>% round(2) %>% format(big.mark = ",")` and a maximum of `r max(training$area, na.rm = TRUE) %>% round(2) %>% format(big.mark = ",")`.

```{r}
summary(training)
```

By viewing the "largest apartments", some things become apparent. First, the are repeated observations on this small sample: see the third and fourth rows, for instance. Second, the first four rows are obviously typos: the seller typed 650,000 instead of 650 squared meters, to cited the first case. Third, the apartment listed on the fifth row is no longer available rising some doubts on its veracity. Last, the sixth "apartment" is actually a house. When analysing the "smallest apartments", everything seems in order.

```{r}
training %>%
  arrange(desc(area)) %>%
  head(10)
```

So, to fix these (1) we eliminate duplicates, (2) we divide by 1,000 the area of those apartments with over 10,000 squared meters of area, (3) remove those apartments that are obviously not of interest.

```{r}
training <- training %>%
  filter(id != "/apartamentos-venta-cuesta-hermosa-ii/1236477/",
         id != "/apartamentos-venta-los-cacicazgos/1272251/") %>%
  mutate(area = ifelse(area > 10000, area / 1000, area)) %>%
  unique()
```

Let's do the same with `price`. Viewing `price` alone might be misleading as a apartment with 30 squared meters could be worth 20,000 dollars, but one with 280 squared meters could hardly be worth $10,256. Price per squared meter could tell us more about how extreme of a value it is.

```{r}
summary(training$price)

training <- training %>%
  mutate(price_per_m2 = price / area,
         area_per_br = area / bedrooms)
```

Anything below $200 seems rather dubious, right? Let's filter them out.But let's use something less subjective.

```{r}
(cutoff <- adjboxStats(training$price_per_m2)$fence)

training <- training %>%
  filter(between(price_per_m2, cutoff[1], cutoff[2]))

summary(training$price_per_m2)
```

Anything below $200 seems rather dubious, right? Let's filter them out.But let's use something less subjective.

```{r}
(cutoff <- adjboxStats(training$area_per_br)$fence)

training <- training %>%
  filter(between(area_per_br, cutoff[1], cutoff[2]))

summary(training$area_per_br)
```

```{r}
glimpse(training)
```

Our data frame is now 20 columns wide and 2,851 rows long.


Many times, the same apartment is listed by different sellers. So, eliminating duplicates is not enough to remove confliting. Some seller are not as rigourous as to list all the amaneties, so we are going to keep the most complete listing:

# Some feature engineering

Now, let's do some feature engineering to help the analysis:

```{r}
glimpse(training)
summary(training)
```


```{r}
training <- training %>%
  na.omit()

nonOutlier <- adjOutlyingness(training)
nonOutlier <- nonOutlier$nonOut

training <- training[nonOutlier, ]

glimpse(training)
summary(training)
```

```{r}
#training <- training[complete.cases(training), ]
#dim(training)
```

Now, our data.frame has `r format(dim(training)[1] %>% round(2), big.mark = ",")` observations and `r format(dim(training)[2] %>% round(2), big.mark = ",")` variables.


# Final removal of outliers

```{r}
trn <- training %>%
  mutate(no_amenities = rowSums(across(planta:gimnasio))) %>%
  arrange(desc(no_amenities)) %>%
  distinct(bathrooms, bedrooms, area, price,
           .keep_all = TRUE)

# Box-Cox tranform
bc <- with(training,
           MASS::boxcox(price ~ area * location))

(lambda <- bc$x[which.max(bc$y)])

new_model <- lm((price ^ lambda - 1)/lambda ~ area, data = training)

ggplot(training, aes(x = (price ^ lambda-1)/lambda)) +
geom_histogram(aes(y = ..density..),
               colour = 1, fill = "white") +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(mean = mean((training$price ^ lambda-1)/lambda),
                            sd = sd((training$price ^ lambda-1)/lambda)),
                col = "#1b98e0")

ggplot(training, aes(x = log(price))) +
  geom_histogram(aes(y = ..density..),
               colour = 1, fill = "white") +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(mean = mean(log(training$price)),
                            sd = sd(log(training$price))),
                col = "#1b98e0")

ggplot(trn, aes(x = log(price))) +
  geom_histogram(aes(y = ..density..),
               colour = 1, fill = "white") +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(mean = mean(log(trn$price)),
                            sd = sd(log(trn$price))),
                col = "#1b98e0")

ggplot(training, aes(x = log(area))) +
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "white") +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(mean = mean(log(training$area)),
                            sd = sd(log(training$area))),
                col = "#1b98e0")

ggplot(trn, aes(x = log(area))) +
  geom_histogram(aes(y = ..density..),
                 colour = 1, fill = "white") +
  geom_density() +
  stat_function(fun = dnorm,
                args = list(mean = mean(log(trn$area)),
                            sd = sd(log(trn$area))),
                col = "#1b98e0")

dec_training <- training %>%
  mutate(decile = cut(log(price), breaks = seq(10, 20, length = 30))) %>%
  group_by(decile) %>%
  summarise(n = n())

dec_trn <- trn %>%
  mutate(decile = cut(log(price), breaks = seq(10, 20, length = 30))) %>%
  group_by(decile) %>%
  summarise(n = n())

dec_training %>%
  left_join(dec_trn, by = "decile") %>%
  mutate(d = n.x - n.y,
         d_per = (n.y / n.x - 1) * 100)

```
```{r}
shapiro.test(log(training$area))
```

# Some more feature engineering

Working on location:

```{r}
#other_loc <- training %>%
#  group_by(location) %>%
#  summarise(n = n()) %>%
#  filter(n < 10) %>%
#  .$location

#training <- training %>%
#  mutate(location = ifelse(location %in% other_loc, "Other", location),
#         location = factor(location))
```

`status` is an ordered categorical variable:

```{r}
#training <- training %>%
#  mutate(status = factor(status, levels = c("En Planos", "En Construcción",
#                                            "Nueva", "Remodelada", "A Remodelar",
#                                            "Segundo Uso", "Fideicomiso")))
```

Transforming other character variables into factor variables:

```{r}

```

